\textbf{SPECFEM3D}\\ \textbf{Cartesian}\\ \textbf{User Manual}

### 2021-11-17 doc/USER_MANUAL/04_creating_databases.tex
renames input parameter subamp_seismos to NTSTEP_BETWEEN_OUTPUT_SAMPLE (making naming more uniform across different SPECFEM codes)

\section{Main parameter file \texttt{Par\_file}}\label{cha:Main-Parameter}

Before running \texttt{xgenerate\_databases}, a number of parameters
need to be set in the main parameter \texttt{Par\_file} located in
the subdirectory \texttt{DATA}:
\begin{description}
\item [{\texttt{SIMULATION\_TYPE}}] is set to 1 for forward simulations,
2 for adjoint simulations (see Section \ref{sec:Adjoint-simulation-finite})
and 3 for kernel simulations (see Section \ref{sec:Finite-Frequency-Kernels}).
\item [{\texttt{SAVE\_FORWARD}}] is only set to \texttt{.true.} for a forward
simulation with the last frame of the simulation saved, as part of
the finite-frequency kernel calculations (see Section \ref{sec:Finite-Frequency-Kernels}).
For a regular forward simulation, leave \texttt{SIMULATION\_TYPE}
and \texttt{SAVE\_FORWARD} at their default values.
\item [{\texttt{UTM\_PROJECTION\_ZONE}}] UTM projection zone in
which your model resides, only valid when \texttt{SUPPRESS\_UTM\_PROJECTION} is \texttt{.false.}.
\item [{\texttt{SUPPRESS\_UTM\_PROJECTION}}] set to be \texttt{.false.}
when your model range is specified in the geographical coordinates,
and needs to be \texttt{.true.} when your model is specified in a
cartesian coordinates. \noun{UTM projection zone in which your simulation
region resides.}
\item [{\texttt{NPROC}}] The number of MPI processors, each one is assigned
one slice of the whole mesh.
\item [{\texttt{NSTEP}}] The number of time steps of the simulation. This
controls the length of the numerical simulation, i.e., twice the number
of time steps requires twice as much CPU time. This feature is not
used at the time of generating the distributed databases but is required
for the solver, i.e., you may change this parameter after running
\texttt{xgenerate\_databases}.
\item [{\texttt{DT}}] The length of each time step in seconds. This feature
is not used at the time of generating the distributed databases but
is required for the solver. Please see also Section~\ref{sec:Choosing-the-Time-Step}
for further details.
\item [{\texttt{NGNOD}}] The number of nodes for 2D and 3D shape functions
for hexahedra. We use either 8-node mesh elements (bricks) or 27-node
elements. If you use the internal mesher, the only option is 8-node
bricks (27-node elements are not supported). \texttt{CUBIT} does not
support HEX27 elements either (it can generate them, but they are
flat, i.e. identical to HEX8). To generate HEX27 elements with curvature
properly taken into account, you can use Gmsh \url{http://geuz.org/gmsh/}
\item [{\texttt{MODEL}}] Must be set to one of the following:

@@

\item [{\texttt{ANISOTROPY}}] Set to \texttt{.true.} if you want to use
an anisotropy model. Please see the file \texttt{model\_aniso.f90}
in subdirectory \texttt{src/generate\_databases/} for the current
implementation of anisotropic models.
\item [{\texttt{TOMOGRAPHY\_PATH}}] Directory in which the tomography files
are stored for using external tomographic Earth models (please read
Chapter \ref{cha:-Changing-the} and Section \ref{sec:Using-tomographic}
`Using external tomographic Earth models' for further details.).
\item [{\texttt{USE\_OLSEN\_ATTENUATION}}] Set to \texttt{.true.} if you
want to use the attenuation model that scaled from the S-wave speed
model using Olsen's empirical relation (see \citet{OlDaBr03}).
\item [{\texttt{OLSEN\_ATTENUATION\_RATIO}}] Determines the Olsen's constant
in Olsen's empirical relation (see \citet{OlDaBr03}).
\item [{\texttt{PML\_CONDITIONS}}] Set to \texttt{.true.} to turn on C-PML
boundary conditions for a regional simulation. Both fluids and elastic solids are supported.
\item [{\texttt{PML\_INSTEAD\_OF\_FREE\_SURFACE}}] Set to \texttt{.true.}
to turn on C-PML boundary conditions on the top surface instead of
the usual free surface.
\item [{\texttt{f0\_FOR\_PML}}] Determines the dominant frequency that
will be used in the calculation of PML damping profiles;
\red{This should be set to the same (or similar) dominant frequency as that of the
source that you will use in your simulation. It is \textbf{VERY IMPORTANT} to do that, otherwise the PML
absorbing conditions can become unstable.}
If you plan to use a Dirac source, then use the dominant frequency of the source wavelet
with which you plan to convolve your seismograms later on in post-processing.
\item [{\texttt{STACEY\_ABSORBING\_CONDITIONS}}] Set to \texttt{.true.}
to turn on Clayton-Enquist absorbing boundary conditions (see \citet{KoTr99}).
In almost all cases it is much better to use CPML absorbing layers
(see the options above) and leave this flag to \texttt{.false.}.
\item [{\texttt{STACEY\_INSTEAD\_OF\_FREE\_SURFACE}}] Set to \texttt{.true.}
to turn on absorbing boundary conditions on the top surface which
by default constitutes a free surface of the model.
\item [{\texttt{BOTTOM\_FREE\_SURFACE}}]
When STACEY\_ABSORBING\_CONDITIONS is set to .true. :
absorbing conditions are defined in xmin, xmax, ymin, ymax and zmin
this option BOTTOM\_FREE\_SURFACE can be set to .true. to
make zmin free surface instead of absorbing condition.

@@

\item [{\texttt{NTSTEP\_BETWEEN\_OUTPUT\_SEISMOS}}] This parameter specifies
the interval at which synthetic seismograms are written in the \texttt{LOCAL\_PATH}
directory. If a run crashes, you may still find usable (but shorter
than requested) seismograms in this directory. On a fast machine set
\texttt{NTSTEP\_BETWEEN\_OUTPUT\_SEISMOS} to a relatively high value
to avoid writing to the seismograms too often. This feature is only
\item [{\texttt{USE\_FORCE\_POINT\_SOURCE}}] Turn this flag on to use a
(tilted) \texttt{FORCESOLUTION} force point source instead of a \texttt{CMTSOLUTION}
moment-tensor source. When the force source does not fall exactly
at a grid point, the solver interpolates the force between grid points
using Lagrange interpolants. This can be useful e.g. for oil industry
foothills simulations in which the source is a vertical force, normal
force, tilted force, or an impact etc. Note that in the \texttt{FORCESOLUTION}
file, you will need to edit the East, North and vertical components
of an arbitrary (not necessarily unitary, the code will normalize it automatically) direction vector of the force vector;
thus refer to Appendix~\ref{cha:Coordinates} for the orientation
of the reference frame. This vector is made unitary internally in
the solver and thus only its direction matters here; its norm is ignored
and the norm of the force used is the factor force source times the
source time function.

When using this option, by default the code can locate the force source
anywhere between mesh points in order to honor its exact location; this
is more precise than using the closest GLL mesh point, but it is also a bit slower.
If needed, you can change that default behavior and force the code to use the closest
GLL mesh point instead by setting flag \texttt{USE\_BEST\_LOCATION} to \texttt{.false.} instead of \texttt{.true.} in file
\texttt{src/shared/constants.h.in} and running the \texttt{configure} script again and recompiling the code.
\item [{\texttt{USE\_RICKER\_TIME\_FUNCTION}}] Turn this flag on to use
a Ricker source time function, i.e., the second derivative of a Gaussian, instead of the source time functions
set by default to represent a (tilted) \texttt{FORCESOLUTION} force
point source or a \texttt{CMTSOLUTION} moment-tensor source.
Note that we use the standard definition of a Ricker, for a dominant frequency $f_0$:
$\mathrm{Ricker}(t) = (1 - 2 a t^2) e^{-a t^2}$, with $a = \pi^2 f_0^2$,
whose Fourier transform is thus:
$\frac{1}{2} \frac{\sqrt{\pi}\omega^2}{a^{3/2}}e^{-\frac{\omega^2}{4 a}}$
This gives the wavelet of Figure~\ref{fig:RickerWavelet}.
%%
\begin{figure}[htbp]
\centering
\includegraphics[width=3in]{figures/Ricker_wavelet.png}
\caption{We use the standard definition of a Ricker (i.e., second derivative of a Gaussian). Image taken from \url{http://subsurfwiki.org}.}
\label{fig:RickerWavelet}
\end{figure}
%%
Originally, if a \texttt{CMTSOLUTION} moment-tensor source is used, a (pseudo)
Heaviside step function with a very short half duration is defined
for elastic cases to represent the permanent slip on the fault while
in the acoustic case a Gaussian source time function with a similarly
short half duration is defined to physically describe actions within
the fluid. Otherwise, if a \texttt{FORCESOLUTION} force source is
used, a (pseudo) Dirac delta source time function is defined by default.
Any other source-time function may then be obtained by convolution.
\item [{\texttt{PRINT\_SOURCE\_TIME\_FUNCTION}}] Turn this flag on to print
information about the source time function in the file \texttt{OUTPUT\_FILES/plot\_source\_time\_function.txt}.
This feature is only relevant for the solver.
%
\item [{\texttt{NUMBER\_OF\_SIMULTANEOUS\_RUNS}}] adds the ability to run several calculations (several earthquakes)
in an embarrassingly-parallel fashion from within the same run;
this can be useful when using a very large supercomputer to compute
many earthquakes in a catalog, in which case it can be better from
a batch job submission point of view to start fewer and much larger jobs,
each of them computing several earthquakes in parallel.

@@

\item [{\texttt{USE\_FAILSAFE\_MECHANISM}}] if one or a few of these simultaneous runs fail,
kill all the runs or let the others finish using a fail-safe mechanism
(in most cases, should be set to true).

@@

\item [\texttt{ADIOS\_ENABLED}] Turn this flag on to enable ADIOS. If set to \texttt{.false.}, subsequent ADIOS
parameters will not be considered.
\item [\texttt{ADIOS\_FOR\_DATABASES}] Turn this flag on to use ADIOS for xmeshfem3D output and
xgenerate\_database input.
\item [\texttt{ADIOS\_FOR\_MESH}]  Turn this flag on to use ADIOS for generated databases.
\item [\texttt{ADIOS\_FOR\_FORWARD\_ARRAYS}] Turn this flag on to read and write forward arrays using ADIOS.
\item [\texttt{ADIOS\_FOR\_KERNELS}] Turn this flag on to produce ADIOS kernels that can later be visualized with the ADIOS version of combine\_vol\_data.

The present version of SPECFEM can handle fully saturated porous simulations, Christina Morency implemented Biot equation.
But the code cannot calculate partially saturated cases in its state.
Christina Morency is presently working on a dual porosity, dual permeability formulation, type Pride and Berryman, for an other project, but it will not be available for some time.\\

The way we prescribe material property for porous material in SPECFEM3D is as follow:
We use a file name "nummaterial\_poroelastic\_file", which is located in the directory \texttt{MESH/}, the format is as follow:
\begin{verbatim}
 rhos rhof phi c kxx kxy kxz kyy kyz kzz Ks Kf Kfr etaf mufr
\end{verbatim}
where\\
\texttt{rho\_s} = solid density,\\
\texttt{rho\_f} = fluid density,\\
\texttt{phi} = porosity,\\
\texttt{tort} = tortuosity,\\
\texttt{kxx} = xx component of permeability tensor,\\
\texttt{kxy} = xy,yx components of permeability tensor,\\
\texttt{kyy} = yy component of permeability tensor,\\
\texttt{kxz} = xz,zx components of permeability tensor,\\
\texttt{kzz} = zz component of permeability tensor,\\
\texttt{kappa\_s} = solid bulk modulus,\\
\texttt{kappa\_f} = fluid bulk modulus,\\
\texttt{kappa\_fr} = frame bulk modulus,\\
\texttt{eta\_f} = fluid viscosity,\\
\texttt{mu\_fr} = frame shear modulus.\\

Using an external mesh (for instance coming from CUBIT/TRELIS), poroelastic materials have the ID number 3, while 1 is acoustic and 2 is elastic (see the example in the package: EXAMPLES/homogeneous\_poroelastic).

@@

If you use PML and an external tomographic velocity and density model,
you should be careful because mathematically a PML cannot handle heterogeneities
along the normal to the PML edge inside the PML layer. This comes
from the fact that the damping profile that is defined assumes a constant
velocity and density model along the normal direction.

Thus, you need to modify your velocity and density model in order
for it to be 1D inside the PML, as shown in Figure~\ref{fig:modify_external_velocity_model_to_use_PML}.

This applies to the bottom layer as well; there you should make sure
that your model is 1D and thus constant along the vertical direction.

>>>>

\newpage
\section{Main parameter file \texttt{Par\_file}}\label{cha:Main-Parameter}

Before running \texttt{xgenerate\_databases}, a number of parameters
need to be set in the main parameter \texttt{Par\_file} located in
the subdirectory \texttt{DATA}:\\

\noindent {\bf $\bullet$ Simulation setup}
\begin{description}
\item [{\texttt{SIMULATION\_TYPE}}] is set to 1 for forward simulations,
2 for adjoint simulations (see Section \ref{sec:Adjoint-simulation-finite})
and 3 for kernel simulations (see Section \ref{sec:Finite-Frequency-Kernels}).
\item [{\texttt{SAVE\_FORWARD}}] is only set to \texttt{.true.} for a forward
simulation with the last frame of the simulation saved, as part of
the finite-frequency kernel calculations (see Section \ref{sec:Finite-Frequency-Kernels}).
For a regular forward simulation, leave \texttt{SIMULATION\_TYPE}
and \texttt{SAVE\_FORWARD} at their default values.
\item [{\texttt{UTM\_PROJECTION\_ZONE}}] UTM projection zone in
which your model resides, only valid when \texttt{SUPPRESS\_UTM\_PROJECTION} is \texttt{.false.}.
\item [{\texttt{SUPPRESS\_UTM\_PROJECTION}}] set to be \texttt{.false.}
when your model range is specified in the geographical coordinates,
and needs to be \texttt{.true.} when your model is specified in a
cartesian coordinates. \noun{UTM projection zone in which your simulation
region resides.}
\item [{\texttt{NPROC}}] The number of MPI processors, each one is assigned
one slice of the whole mesh.
\item [{\texttt{NSTEP}}] The number of time steps of the simulation. This
controls the length of the numerical simulation, i.e., twice the number
of time steps requires twice as much CPU time. This feature is not
used at the time of generating the distributed databases but is required
for the solver, i.e., you may change this parameter after running
\texttt{xgenerate\_databases}.
\item [{\texttt{DT}}] The length of each time step in seconds. This feature
is not used at the time of generating the distributed databases but
is required for the solver. Please see also Section~\ref{sec:Choosing-the-Time-Step}
for further details.
\end{description}

\vspace{1cm}
\noindent {\bf $\bullet$ Mesh setup}
\begin{description}
\item [{\texttt{NGNOD}}] The number of nodes for 2D and 3D shape functions
for hexahedra. We use either 8-node mesh elements (bricks) or 27-node
elements. If you use the internal mesher, the only option is 8-node
bricks (27-node elements are not supported). \texttt{CUBIT} does not
support HEX27 elements either (it can generate them, but they are
flat, i.e. identical to HEX8). To generate HEX27 elements with curvature
properly taken into account, you can use Gmsh \url{http://geuz.org/gmsh/}
\item [{\texttt{MODEL}}] Must be set to one of the following:

@@

\item [{\texttt{ANISOTROPY}}] Set to \texttt{.true.} if you want to use
an anisotropy model. Please see the file \texttt{model\_aniso.f90}
in subdirectory \texttt{src/generate\_databases/} for the current
implementation of anisotropic models.
\item [{\texttt{TOMOGRAPHY\_PATH}}] Directory in which the tomography files
are stored for using external tomographic Earth models (please read
Chapter \ref{cha:-Changing-the} and Section \ref{sec:Using-tomographic}
`Using external tomographic Earth models' for further details.).
\item [{\texttt{USE\_OLSEN\_ATTENUATION}}] Set to \texttt{.true.} if you
want to use the attenuation model that scaled from the S-wave speed
model using Olsen's empirical relation (see \citet{OlDaBr03}).
\item [{\texttt{OLSEN\_ATTENUATION\_RATIO}}] Determines the Olsen's constant
in Olsen's empirical relation (see \citet{OlDaBr03}).
\end{description}

\vspace{1cm}
\noindent {\bf $\bullet$ Absorbing boundary conditions setup}
\begin{description}
\item [{\texttt{PML\_CONDITIONS}}] Set to \texttt{.true.} to turn on C-PML
boundary conditions for a regional simulation. Both fluids and elastic solids are supported.
\item [{\texttt{PML\_INSTEAD\_OF\_FREE\_SURFACE}}] Set to \texttt{.true.}
to turn on C-PML boundary conditions on the top surface instead of
the usual free surface.
\item [{\texttt{f0\_FOR\_PML}}] Determines the dominant frequency that
will be used in the calculation of PML damping profiles;
\red{This should be set to the same (or similar) dominant frequency as that of the
source that you will use in your simulation. It is \textbf{VERY IMPORTANT} to do that, otherwise the PML
absorbing conditions can become unstable.}
If you plan to use a Dirac source, then use the dominant frequency of the source wavelet
with which you plan to convolve your seismograms later on in post-processing.
\item [{\texttt{STACEY\_ABSORBING\_CONDITIONS}}] Set to \texttt{.true.}
to turn on Clayton-Enquist absorbing boundary conditions (see \citet{KoTr99}).
In almost all cases it is much better to use CPML absorbing layers
(see the options above) and leave this flag to \texttt{.false.}.
\item [{\texttt{STACEY\_INSTEAD\_OF\_FREE\_SURFACE}}] Set to \texttt{.true.}
to turn on absorbing boundary conditions on the top surface which
by default constitutes a free surface of the model.
\item [{\texttt{BOTTOM\_FREE\_SURFACE}}]
When STACEY\_ABSORBING\_CONDITIONS is set to .true. :
absorbing conditions are defined in xmin, xmax, ymin, ymax and zmin
this option BOTTOM\_FREE\_SURFACE can be set to .true. to
make zmin free surface instead of absorbing condition.
\end{description}

\vspace{1cm}
\noindent {\bf $\bullet$ Visualization setup}
\begin{description}

@@

\vspace{1cm}
\noindent {\bf $\bullet$ Sources setup}
\begin{description}
\item [{\texttt{USE\_FORCE\_POINT\_SOURCE}}] Turn this flag on to use a
(tilted) \texttt{FORCESOLUTION} force point source instead of a \texttt{CMTSOLUTION}
moment-tensor source. When the force source does not fall exactly
at a grid point, the solver interpolates the force between grid points
using Lagrange interpolants. This can be useful e.g. for oil industry
foothills simulations in which the source is a vertical force, normal
force, tilted force, or an impact etc. Note that in the \texttt{FORCESOLUTION}
file, you will need to edit the East, North and vertical components
of an arbitrary (not necessarily unitary, the code will normalize it automatically) direction vector of the force vector;
thus refer to Appendix~\ref{cha:Coordinates} for the orientation
of the reference frame. This vector is made unitary internally in
the solver and thus only its direction matters here; its norm is ignored
and the norm of the force used is the factor force source times the
source time function.

When using this option, by default the code can locate the force source
anywhere between mesh points in order to honor its exact location; this
is more precise than using the closest GLL mesh point, but it is also a bit slower.
If needed, you can change that default behavior and force the code to use the closest
GLL mesh point instead by setting flag \texttt{USE\_BEST\_LOCATION} to \texttt{.false.} instead of \texttt{.true.} in file
\texttt{src/shared/constants.h.in} and running the \texttt{configure} script again and recompiling the code.
\item [{\texttt{USE\_RICKER\_TIME\_FUNCTION}}] Turn this flag on to use
a Ricker source time function, i.e., the second derivative of a Gaussian, instead of the source time functions
set by default to represent a (tilted) \texttt{FORCESOLUTION} force
point source or a \texttt{CMTSOLUTION} moment-tensor source.
Note that we use the standard definition of a Ricker, for a dominant frequency $f_0$:
$\mathrm{Ricker}(t) = (1 - 2 a t^2) e^{-a t^2}$, with $a = \pi^2 f_0^2$,
whose Fourier transform is thus:
$\frac{1}{2} \frac{\sqrt{\pi}\omega^2}{a^{3/2}}e^{-\frac{\omega^2}{4 a}}$
This gives the wavelet of Figure~\ref{fig:RickerWavelet}.
%%
\begin{figure}[htbp]
\centering
\includegraphics[width=3in]{figures/Ricker_wavelet.png}
\caption{We use the standard definition of a Ricker (i.e., second derivative of a Gaussian). Image taken from \url{http://subsurfwiki.org}.}
\label{fig:RickerWavelet}
\end{figure}
%%
Originally, if a \texttt{CMTSOLUTION} moment-tensor source is used, a (pseudo)
Heaviside step function with a very short half duration is defined
for elastic cases to represent the permanent slip on the fault while
in the acoustic case a Gaussian source time function with a similarly
short half duration is defined to physically describe actions within
the fluid. Otherwise, if a \texttt{FORCESOLUTION} force source is
used, a (pseudo) Dirac delta source time function is defined by default.
Any other source-time function may then be obtained by convolution.
\item [{\texttt{PRINT\_SOURCE\_TIME\_FUNCTION}}] Turn this flag on to print
information about the source time function in the file \texttt{OUTPUT\_FILES/plot\_source\_time\_function.txt}.
This feature is only relevant for the solver.
%
\end{description}

\vspace{1cm}
\noindent {\bf $\bullet$ Additionals}
\begin{description}
\item [{\texttt{NTSTEP\_BETWEEN\_OUTPUT\_SEISMOS}}] This parameter specifies
the interval at which synthetic seismograms are written in the \texttt{LOCAL\_PATH}
directory. If a run crashes, you may still find usable (but shorter
than requested) seismograms in this directory. On a fast machine set
\texttt{NTSTEP\_BETWEEN\_OUTPUT\_SEISMOS} to a relatively high value
to avoid writing to the seismograms too often. This feature is only
relevant for the solver.

\item [{\texttt{NUMBER\_OF\_SIMULTANEOUS\_RUNS}}] adds the ability to run several calculations (several earthquakes)
in an embarrassingly-parallel fashion from within the same run;
this can be useful when using a very large supercomputer to compute
many earthquakes in a catalog, in which case it can be better from
a batch job submission point of view to start fewer and much larger jobs,
each of them computing several earthquakes in parallel.

@@

%\item [{\texttt{USE\_FAILSAFE\_MECHANISM}}] if one or a few of these simultaneous runs fail,
%kill all the runs or let the others finish using a fail-safe mechanism
%(in most cases, should be set to true).

@@

\item [\texttt{ADIOS\_ENABLED}] Turn this flag on to enable ADIOS. If set to \texttt{.false.}, subsequent ADIOS
parameters will not be considered.
\item [\texttt{ADIOS\_FOR\_DATABASES}] Turn this flag on to use ADIOS for xmeshfem3D output and
xgenerate\_database input.
\item [\texttt{ADIOS\_FOR\_MESH}]  Turn this flag on to use ADIOS for generated databases.
\item [\texttt{ADIOS\_FOR\_FORWARD\_ARRAYS}] Turn this flag on to read and write forward arrays using ADIOS.
\item [\texttt{ADIOS\_FOR\_KERNELS}] Turn this flag on to produce ADIOS kernels that can later be visualized with the ADIOS version of combine\_vol\_data.
\end{description}
There are quite a few more parameters to change the default setup of your runs. Please check the comments in the \texttt{Par\_file} directly for further explanations.\\

The present version of SPECFEM can handle fully saturated porous simulations.
(Christina Morency implemented Biot's equation.
However, the code cannot calculate partially saturated cases in its state.
Christina Morency is presently working on a dual porosity, dual permeability formulation, type Pride and Berryman, for an other project, but it will not be available for some time.)
The way we prescribe material property for porous material in SPECFEM3D is as follow.
We use a file \texttt{nummaterial\_poroelastic\_file}, which is located in the directory \texttt{MESH/}. The line format is as follow:
{\small \begin{verbatim}
  rhos rhof phi c kxx kxy kxz kyy kyz kzz Ks Kf Kfr etaf mufr
\end{verbatim}}
\begin{description}
\item where\\
\texttt{rho\_s} = solid density,\\
\texttt{rho\_f} = fluid density,\\
\texttt{phi} = porosity,\\
\texttt{tort} = tortuosity,\\
\texttt{kxx} = xx component of permeability tensor,\\
\texttt{kxy} = xy,yx components of permeability tensor,\\
\texttt{kyy} = yy component of permeability tensor,\\
\texttt{kxz} = xz,zx components of permeability tensor,\\
\texttt{kzz} = zz component of permeability tensor,\\
\texttt{kappa\_s} = solid bulk modulus,\\
\texttt{kappa\_f} = fluid bulk modulus,\\
\texttt{kappa\_fr} = frame bulk modulus,\\
\texttt{eta\_f} = fluid viscosity,\\
\texttt{mu\_fr} = frame shear modulus.\\
\end{description}
Using an external mesh (for instance coming from CUBIT/TRELIS), poroelastic materials have the ID number 3, while 1 is acoustic and 2 is elastic (see the example in the package: EXAMPLES/homogeneous\_poroelastic).\\

@@

If you use PML and an external tomographic velocity and density model,
you should be careful because mathematically a PML cannot handle heterogeneities
along the normal to the PML edge inside the PML layer. This comes
from the fact that the damping profile that is defined assumes a constant
velocity and density model along the normal direction.
Thus, you need to modify your velocity and density model in order
for it to be 1D inside the PML, as shown in Figure~\ref{fig:modify_external_velocity_model_to_use_PML}.
This applies to the bottom layer as well; there you should make sure
that your model is 1D and thus constant along the vertical direction.

### 2021-11-17 doc/USER_MANUAL/05_running_the_solver.tex
renames input parameter subamp_seismos to NTSTEP_BETWEEN_OUTPUT_SAMPLE (making naming more uniform across different SPECFEM codes)

\noindent
The solver needs three input files in the \texttt{DATA} directory
to run:
\begin{description}
\item [{\texttt{Par\_file}}] the main parameter file which was discussed
in detail in the previous Chapter~\ref{cha:Creating-Distributed-Databases},
\item [{\texttt{CMTSOLUTION}} or {\texttt{FORCESOLUTION}}] the earthquake
source parameter file or the force source parameter file, and
\item [{\texttt{STATIONS}}] the stations file.
\end{description}
Most parameters in the \texttt{Par\_file} should be set prior to running
the databases generation. Only the following parameters may be changed
after running \texttt{xgenerate\_databases}:
\begin{itemize}
\item the simulation type control parameters: \texttt{SIMULATION\_TYPE}
and \texttt{SAVE\_FORWARD}
\item the time step parameters \texttt{NSTEP} and \texttt{DT}
\item the absorbing boundary control parameter \texttt{PML\_CONDITIONS}
on condition that the\\
 \texttt{PML\_INSTEAD\_OF\_FREE\_SURFACE} flag remains unmodified
after running the databases generation.
\item the movie control parameters \texttt{MOVIE\_SURFACE}, \texttt{MOVIE\_VOLUME},
and \texttt{NTSTEPS\_BETWEEN\_FRAMES}
\item the ShakeMap\textregistered{} option \texttt{CREATE\_SHAKEMAP}
\item the output information parameters \texttt{MOVIE\_TYPE}, \texttt{NTSTEP\_BETWEEN\_OUTPUT\_INFO} and\\
\texttt{NTSTEP\_BETWEEN\_OUTPUT\_SEISMOS}
\item the \texttt{PRINT\_SOURCE\_TIME\_FUNCTION} flags
\end{itemize}
Any other change to the \texttt{Par\_file} implies rerunning both
the database generator \texttt{xgenerate\_databases} and the solver
\texttt{xspecfem3D}.

@@

Each line represents one station in the following format:
{\small
\begin{verbatim}
Station Network Latitude(degrees) Longitude(degrees) Elevation(m) burial(m)
\end{verbatim}
}
The solver \texttt{xspecfem3D} filters the list of stations in file
\texttt{DATA/STATIONS} to exclude stations that are not located within
the region given in the \texttt{Par\_file} (between \texttt{LATITUDE\_MIN}
and \texttt{LATITUDE\_MAX} and between \texttt{LONGITUDE\_MIN} and
\texttt{LONGITUDE\_MAX}). The filtered file is called \texttt{DATA/STATIONS\_FILTERED}.

Elevation and burial are generally applicable to geographical regions.
Burial is measured down from the top surface. For other problems in other
fields (ultrasonic testing, medical imaging etc...),
it may be confusing. We generally follow either of the following
procedures for those kind of problems:

Procedure 1 (mostly for geophysics, when the top surface is a free surface (topography) and the five other edges of the mesh are absorbing surfaces):\\
====================================================================================================================================================\\

- Put the origin on the top of the model.\\

- Let's say you want to place two receivers at (x1,y1,z1) and
(x2,y2,z2). Your STATIONS file should look like:\\

BONE  GR  y1  x1  0.00  -z1\\
BONE  GR  y2  x2  0.00  -z2\\

Procedure 2 (useful for other application domains, in which using the absolute $Z$ position of the sources and receivers is more standard than using their depth from the surface):\\
==================================================================================================================================================================================

- In principle in the international CMTSOLUTION format in geophysics
the depth is given in kilometers; however for users in other fields (non-destructive testing, medical imaging, near-surface studies...) who may prefer to give the position of the source (rather than its depth from the surface), or for people who use FORCESOLUTION to describe the source rather than CMTSOLUTION, we provide an option called USE\_SOURCES\_RECEIVERS\_Z in the Par\_file, and if so that position is read from CMTSOLUTION in meters rather than kilometers (and again, it is then the true position in the mesh, not the depth). When option USE\_SOURCES\_RECEIVERS\_Z in the Par\_file is on, this remark applies to the position of the receivers as well.\\

- Let's say you want to place two receivers at (x1,y1,z1) and
(x2,y2,z2). Your STATIONS file should then look like:\\
\\
BONE  GR  y1  x1  0.00  z1\\
BONE  GR  y2  x2  0.00  z2\\
\\
The option USE\_SOURCES\_RECEIVERS\_Z set to .true. will then discard the elevation and set
burial as the $z$ coordinate.\\
\\
Third column is Y and Fourth is X due to the latitude/longitude convention.\\
\\
You can replace the station name "BONE" with any word of length less
than 32, and the network name "GR" with any word of length less than
8.\\
\\
You can always plot OUTPUT\_FILES/sr.vtk file in ParaView to check the
source/receiver locations after your simulation.\\

>>>>

\noindent
The solver needs three input files in the \texttt{DATA} directory
to run:
\begin{itemize}
\item {\bf \texttt{Par\_file}} the main parameter file which was discussed
in detail in the previous Chapter~\ref{cha:Creating-Distributed-Databases},
\item {\bf \texttt{CMTSOLUTION}} or {\bf \texttt{FORCESOLUTION}} the earthquake
source parameter file or the force source parameter file, and
\item {\bf \texttt{STATIONS}} the stations file.
\end{itemize}

\noindent
Most parameters in the \texttt{Par\_file} should be set prior to running
the databases generation. Only the following parameters may be changed
after running \texttt{xgenerate\_databases}:
\begin{itemize}
\item the simulation type control parameters \texttt{SIMULATION\_TYPE}
and \texttt{SAVE\_FORWARD}
\item the time step parameters \texttt{NSTEP} and \texttt{DT}
\item the absorbing boundary control parameter \texttt{PML\_CONDITIONS}
on condition that the\\
 \texttt{PML\_INSTEAD\_OF\_FREE\_SURFACE} flag remains unmodified
after running the databases generation.
\item the movie control parameters \texttt{MOVIE\_SURFACE}, \texttt{MOVIE\_VOLUME},
and \texttt{NTSTEPS\_BETWEEN\_FRAMES}
\item the ShakeMap\textregistered{} option \texttt{CREATE\_SHAKEMAP}
\item the output information parameters \texttt{MOVIE\_TYPE}, \texttt{NTSTEP\_BETWEEN\_OUTPUT\_INFO} and\\
\texttt{NTSTEP\_BETWEEN\_OUTPUT\_SEISMOS}
\item the \texttt{PRINT\_SOURCE\_TIME\_FUNCTION} flags
\end{itemize}
Any other change to the \texttt{Par\_file} implies rerunning both
the database generator \texttt{xgenerate\_databases} and the solver
\texttt{xspecfem3D}.\\

@@

\noindent
Each line represents one station in the following format:
{\small
\begin{verbatim}
Station Network Latitude(degrees) Longitude(degrees) Elevation(m) burial(m)
\end{verbatim}
}
The solver \texttt{xspecfem3D} filters the list of stations in file
\texttt{DATA/STATIONS} to exclude stations that are not located within
the region given in the \texttt{Par\_file} (between \texttt{LATITUDE\_MIN}
and \texttt{LATITUDE\_MAX} and between \texttt{LONGITUDE\_MIN} and
\texttt{LONGITUDE\_MAX}). The filtered file is called \texttt{DATA/STATIONS\_FILTERED}.
Elevation and burial are generally applicable to geographical regions.
Burial is measured down from the top surface.\\

\noindent
For other problems in other
fields (ultrasonic testing, medical imaging etc...),
it may be confusing. We generally follow either one of the following
procedures for those kind of problems:
\begin{itemize}
\item {\it Procedure 1:} mostly for geophysics, when the top surface is a free surface (topography) and the five other edges of the mesh are absorbing surfaces
\begin{enumerate}
\item [-] Put the origin on the top of the model.\\
\item [-] Let's say you want to place two receivers at (x1,y1,z1) and
(x2,y2,z2). Your STATIONS file should look like:
{\small \begin{verbatim}
BONE  GR  y1  x1  0.00  -z1
BONE  GR  y2  x2  0.00  -z2
\end{verbatim}}
\end{enumerate}

\item {\it Procedure 2:} useful for other application domains, in which using the absolute $Z$ position of the sources and receivers is more standard than using their depth from the surface
\begin{enumerate}
\item [-] In principle in the international CMTSOLUTION format in geophysics
the depth is given in kilometers; however for users in other fields (non-destructive testing, medical imaging, near-surface studies...) who may prefer to give the position of the source (rather than its depth from the surface), or for people who use FORCESOLUTION to describe the source rather than CMTSOLUTION, we provide an option called USE\_SOURCES\_RECEIVERS\_Z in the Par\_file, and if so that position is read from CMTSOLUTION in meters rather than kilometers (and again, it is then the true position in the mesh, not the depth). When option USE\_SOURCES\_RECEIVERS\_Z in the Par\_file is on, this remark applies to the position of the receivers as well.\\
\item [-] Let's say you want to place two receivers at (x1,y1,z1) and
(x2,y2,z2). Your STATIONS file should then look like:
{\small \begin{verbatim}
BONE  GR  y1  x1  0.00  z1
BONE  GR  y2  x2  0.00  z2
\end{verbatim}}
The option USE\_SOURCES\_RECEIVERS\_Z set to .true. will then discard the elevation and set
burial as the $z$ coordinate.
Third column is Y and Fourth is X due to the latitude/longitude convention.\\
\end{enumerate}

You can replace the station name "BONE" with any word of length less
than 32, and the network name "GR" with any word of length less than
8.
And you can always plot OUTPUT\_FILES/sr.vtk file in ParaView to check the
source/receiver locations after your simulation.\\
\end{itemize}
